Smoking and COVID Living Review (v7): a bayesian analysis

### 2020-09-24 reports/bayesian_report.Rmd
preparing v8

```{r load_scripts}
source(here("scripts", "author_dictionary.R"))
```
## Download data
We maintain a Google sheets document [here](sheets_id) which contains all eligible articles the research team have identified. This is currently viewable only and not open for editing. We have collected the general information for the search in the first sheet, the studies general information in the second, further information on study populations, outcomes of interest and quality of the studies are contained in the further sheets. These sheets are organised in a similar way to the manuscript and so are labelled table_1 to table_6.

``` {r google_data}
sheets_id <- as_sheets_id("https://docs.google.com/spreadsheets/d/1HV2Zk_UD2_s1zzd47dCyb4UKRSM0dDg7zmUPA9CMdx4/edit?usp=sharing")
search_details <- read_sheet(sheets_id, range = 'article_screening')
data_study_general <- read_sheet(sheets_id, range =  'general_details')
table_1 <- read_sheet(sheets_id, sheet = 'pop_descriptives')
table_2 <-  read_sheet(sheets_id, sheet = 'testing')
table_3 <-  read_sheet(sheets_id, sheet = 'hospitalisation')
table_4 <-  read_sheet(sheets_id, sheet = 'severity')
table_5 <-  read_sheet(sheets_id, sheet = 'mortality')
table_6 <-  read_sheet(sheets_id, sheet = 'updated_quality_appraisal')
```
We have extracted data from `r nrow(data_study_general)` studies, however, some of these are duplicates and some have been removed from the analysis set. For example there have been multiple versions of the ISARIC report and several publications have been retracted. We aim to keep on top of this but there is some risk that in between versions data may be superseded.

@@

```{r head_search_details}
flextable(search_details[,1:11])
```
The second sheet is **General study details** this includes the lead author, where the study is published, the date it was published, the country in which the data originated and the DoI or link to the publication for our records.
This is a searchable list of all the studies that have met out inclusion and exclusion criteria up until `r paste(max(search_details$date_screened))`. These are live links and should direct to the relevant article, however, some may have been deleted or require institutional access.
```{r tail_study_details, include=FALSE}
datatable(data_study_general[1:4] %>%
            mutate(`date_published_format(2020-05-12)` = as.Date(data_study_general$`date_published_format(2020-05-12)`),
                   doi = paste0("<a href='",doi,"' target='_blank'>",doi,"</a>")),
          colnames = c("Lead author" = 2, "Date published" = 3, "Source" = 4, "DOI" = 5),
          escape = F)
```
The third sheet produces **Table 1** this contains the details of the population of the studies in our review. Primarily includes smoking status but also the proportion of females in the study and the age of participants.
```{r tail_table_1, include=FALSE}
datatable(table_1 %>%
                 filter(lead_author != is.na(lead_author)))
```
The fourth sheet produces **Table 2** this contains the data extracted from relevant studies on testing data and smoking status. Most studies will not have data on this outcome however we filter these subsequently. These are sparse datasets as many studies do not report on all levels of smoking exposure.
```{r tail_table_2, include=FALSE}
datatable(
  table_2 %>%
  filter(data_on_testing == TRUE)
)
```
The fifth sheet produces **Table 3** this contains the data extracted from relevant studies on hospitalisation and smoking status. 
```{r tail_table_3, include=FALSE}
datatable(
  table_3 %>%
  filter(data_on_hospitalisation == TRUE)
)
```
The sixth sheet produces **Table 4** this contains the data extracted from relevant studies disease severity data and smoking status. Disease severity is differentially reported across the studies in some there is reference to WHO severity score in others we have based the distinction on clinical characteristics or whether patients were admitted to Intensive or Critical care areas.
```{r tail_table_4, include=FALSE}
datatable(
  table_4 %>%
  filter(data_disease_severity == TRUE)
)
```
The seventh sheet produces **Table 5** this contains the data extracted from relevant studies on COVID-19 related mortality and smoking status. A caution needed when interpreting this data is that most studies have not followed patients up for sufficient amounts of time to be able to accurately report on disease associated mortality.
```{r tail_table_5, include=FALSE}
datatable(
  table_5 %>%
  filter(data_on_deaths == TRUE)
)
```
The final sheet is used to collate data on **Study quality** we changed our method on assessing study quality in version 3 of the review. We use the following criteria:

@@

prev_versions <- c('v1', 'v2', 'v3', 'v4', 'v5', 'v6')
#This can categorise which studies we want to look at
current_version <- c('v7')
#This will categorise which studies we are including in the current report

analysed_versions <- c('v1', 'v2', 'v3', 'v4', 'v5', 'v6', 'v7')
#This will incorporate all studies into the current version of the report
```
We define the current version of the report here. At the moment this needs to be changed directly when compiling the code. This report has been generated for version `r paste(current_version)`

``` {r excluding_studies}
exclude_from_analysis <- c('isaric_1', 'isaric_2', 'isaric_3', 'miyara_old', 'isaric_4', 'mehra', 'hopkinson', 'gaibazzi', 'miyara_updated', 'russell_old', "petrilli_old", "magagnoli_old", "niedzwiedz_old", "bello_chavolla_old", "kimmig_old", "russell_old", "zuo_zuo_old", "patel_old", "shi_zuo_old")
```
We don't want to remove studies from the dataset despite them no longer being included in the analysis. The excluded studies are defined in this chunk of code. We have excluded this version the following studies:
-`r knitr::combine_words(exclude_from_analysis)`
Generally we label these as _old in the Google sheet but there are exceptions to this so we manually define them at the moment.

@@

my_bins <- c(0, 1, 5, 10, 20, 30, 40, 70)
mypalette <- colorBin(palette="YlOrBr", domain=my_spdf@data$Number, na.color="transparent", bins=my_bins)

@@

#Smoking prevalence by country
country_prevalence_list <- table_1 %>%
    group_by(country, add = T) %>%
  mutate(., current_smok_percentage = current_smoker/total*100) %>%
  mutate(., former_smok_percentage = former_smoker/total*100) %>%
  mutate(., missing_percentage = missing/total*100) %>%
  mutate(., not_stated_percentage = not_stated/total*100) %>%
  mutate(never_smoker_percentage = never_smoker/total*100) %>%
  group_split()

@@

save_as_docx(a, path = here('output_data', 'Table_1.docx'))
```

@@

included_studies <- quality_table_2 %>%
  filter(overall_quality != 'poor') %>%
  select(lead_author) %>%
  filter(lead_author != "reiter")

meta <- tibble('author' = table_2$lead_author,
               'negative_smoker' = table_2$negative_current_smoker,
               'negative_never_smoker' = table_2$negative_never_smoker,
               'positive_smoker' = table_2$positive_current_smoker,
               'positive_never_smoker' = table_2$positive_never_smoker,
               'negative_former_smoker' = table_2$negative_former_smoker,
               'positive_former_smoker' = table_2$positive_former_smoker) %>%
        filter(author %in% included_studies$lead_author)
meta$author <- to_upper_camel_case(meta$author, sep_out = ", ")
meta$author <- plyr::mapvalues(meta$author,
                               from = cleaned_names,
                               to = correct_names)

testing_rr <- list()
testing_rr[[1]] <- RR_testing('Rentsch', 'current')
testing_rr[[2]] <- RR_testing('Rentsch', 'former')
testing_rr[[3]] <- RR_testing('Cho', 'current')
testing_rr[[4]] <- RR_testing('Cho', 'former')
testing_rr[[5]] <- RR_testing('Kolin', 'current')
testing_rr[[6]] <- RR_testing('Kolin', 'former')
testing_rr[[7]] <- RR_testing('de Lusignan', 'current')
testing_rr[[8]] <- RR_testing('de Lusignan', 'former')
testing_rr[[9]] <- RR_testing('Parrotta', 'current')
testing_rr[[10]] <- RR_testing('Parrotta', 'former')
testing_rr[[11]] <- RR_testing('Israel', 'current')
testing_rr[[12]] <- RR_testing('Israel', 'former')
testing_rr[[13]] <- RR_testing('Eugen-Olsen', 'current')
testing_rr[[14]] <- RR_testing('Eugen-Olsen', 'former')
testing_rr[[15]] <- RR_testing('Houlihan', 'current')
testing_rr[[16]] <- RR_testing('Houlihan', 'former')
testing_rr[[17]] <- RR_testing('Woolford', 'current')
testing_rr[[18]] <- RR_testing('Woolford', 'former')
testing_rr[[19]] <- RR_testing('Govind', 'current')
testing_rr[[20]] <- RR_testing('Govind', 'former')
testing_rr[[21]] <- RR_testing('Gu', 'current')
testing_rr[[22]] <- RR_testing('Gu', 'former')
testing_rr[[23]] <- RR_testing('Petrilli', 'current')
testing_rr[[24]] <- RR_testing('Petrilli', 'former')
testing_rr[[25]] <- RR_testing('Auvinen', 'current')
testing_rr[[26]] <- RR_testing('Auvinen', 'former')
testing_rr[[27]] <- RR_testing('Merkely', 'current')
testing_rr[[28]] <- RR_testing('Merkely', 'former')
testing_rr[[29]] <- RR_testing('Iversen', 'current')
testing_rr[[30]] <- RR_testing('Iversen', 'former')
testing_rr[[31]] <- RR_testing('Fillmore', 'current')
testing_rr[[32]] <- RR_testing('Fillmore', 'former')

data <- testing_rr
k <- do.call(rbind.data.frame, data)

@@

#### Current versus never smoking for risk of severe disease
``` {r}
# Data --------------------------------------------------------------------
meta <- tibble('author' = table_4$lead_author,
               'non_severe_smoker' = table_4$non_severe_current_smoker,
               'non_severe_never_smoker' = table_4$non_severe_never_smoker,
               'severe_smoker' = table_4$severe_disease_current_smoker,
               'severe_never_smoker' = table_4$severe_disease_never_smoker,
               'non_severe_former_smoker' = table_4$non_severe_former_smoker,
               'severe_former_smoker' = table_4$severe_disease_former_smoker) %>%
  filter(author %in% included_studies$lead_author)

>>>>

```{r load_scripts}
source(here("scripts", "author_dictionary.R"))
```
## Download data
``` {r google_data}
sheets_id <- as_sheets_id("https://docs.google.com/spreadsheets/d/1sTlqMbGiWDWJT3dTpa3YQU4z6qcWb0v6Wkp40RcWLXQ/edit?usp=sharing")
search_details <- read_sheet(sheets_id, range = 'article_screening')
data_study_general <- read_sheet(sheets_id, range =  'general_details')
table_1 <- read_sheet(sheets_id, sheet = 'pop_descriptives')
table_2 <-  read_sheet(sheets_id, sheet = 'testing')
table_3 <-  read_sheet(sheets_id, sheet = 'hospitalisation')
table_4 <-  read_sheet(sheets_id, sheet = 'severity')
table_5 <-  read_sheet(sheets_id, sheet = 'mortality')
table_6 <-  read_sheet(sheets_id, sheet = 'updated_quality_appraisal')
```
We maintain a Google sheets document [here](sheets_id) which contains all eligible articles the research team have identified. This is currently viewable only and not open for editing. We have collected the general information for the search in the first sheet, the studies general information in the second, further information on study populations, outcomes of interest and quality of the studies are contained in the further sheets. These sheets are organised in a similar way to the manuscript and so are labelled table_1 to table_6.

The pull requests from the API may happen too rapidly if this is run as a chunk so it's better to run this line by line.

We have extracted data from `r nrow(data_study_general)` studies, however, some of these are duplicates and some have been removed from the analysis set. For example there have been multiple versions of the ISARIC report and several publications have been retracted. We aim to keep on top of this but there is some risk that in between versions data may be superseded.

@@

```{r head_search_details}
flextable(search_details[,1:12])
```
The second sheet is **General study details** this includes the lead author, where the study is published, the date it was published, the country in which the data originated and the DoI or link to the publication for our records.
This is a searchable list of all the studies that have met out inclusion and exclusion criteria up until `r paste(max(search_details$date_screened))`. These are live links and should direct to the relevant article, however, some may have been deleted or require institutional access.
```{r tail_study_details, include=FALSE}
datatable(data_study_general[1:4] %>%
            mutate(`date_published_format(2020-05-12)` = as.Date(data_study_general$`date_published_format(2020-05-12)`),
                   doi = paste0("<a href='",doi,"' target='_blank'>",doi,"</a>")),
          colnames = c("Lead author" = 2, "Date published" = 3, "Source" = 4, "DOI" = 5),
          escape = F)
```
The third sheet produces **Table 1** this contains the details of the population of the studies in our review. Primarily includes smoking status but also the proportion of females in the study and the age of participants.
```{r tail_table_1, include=FALSE}
datatable(table_1 %>%
                 filter(lead_author != is.na(lead_author)))
```
The fourth sheet produces **Table 2** this contains the data extracted from relevant studies on testing data and smoking status. Most studies will not have data on this outcome however we filter these subsequently. These are sparse datasets as many studies do not report on all levels of smoking exposure.
```{r tail_table_2, include=FALSE}
datatable(
  table_2 %>%
  filter(data_on_testing == TRUE)
)
```
The fifth sheet produces **Table 3** this contains the data extracted from relevant studies on hospitalisation and smoking status. 
```{r tail_table_3, include=FALSE}
datatable(
  table_3 %>%
  filter(data_on_hospitalisation == TRUE)
)
```
The sixth sheet produces **Table 4** this contains the data extracted from relevant studies disease severity data and smoking status. Disease severity is differentially reported across the studies in some there is reference to WHO severity score in others we have based the distinction on clinical characteristics or whether patients were admitted to Intensive or Critical care areas.
```{r tail_table_4, include=FALSE}
datatable(
  table_4 %>%
  filter(data_disease_severity == TRUE)
)
```
The seventh sheet produces **Table 5** this contains the data extracted from relevant studies on COVID-19 related mortality and smoking status. A caution needed when interpreting this data is that most studies have not followed patients up for sufficient amounts of time to be able to accurately report on disease associated mortality.
```{r tail_table_5, include=FALSE}
datatable(
  table_5 %>%
  filter(data_on_deaths == TRUE)
)
```
The final sheet is used to collate data on **Study quality** we changed our method on assessing study quality in version 3 of the review. We use the following criteria:

@@

prev_versions <- head(unique(data_study_general$review_version), n = -1)
#This can categorise which studies we want to look at
current_version <- tail(unique(data_study_general$review_version), n = 1)
#This will categorise which studies we are including in the current report

analysed_versions <- c(prev_versions, current_version)
#This will incorporate all studies into the current version of the report
```
We define the current version of the report here. This report has been generated for version `r paste(current_version)`

``` {r excluding_studies}
superseded_studies <- names(tidyselect::vars_select(data_study_general$lead_author, ends_with("_old", ignore.case = T)))
old_isaric <- head(names(tidyselect::vars_select(data_study_general$lead_author, starts_with("isaric", ignore.case = T))), n = -1)
excluded_other_reason <- c("mehra", #retracted
                           "hopkinson", #self-reported
                           "gaibazzi" #withdrawn
                           )

exclude_from_analysis <- c(superseded_studies, old_isaric, excluded_other_reason)
```
We don't want to remove studies from the dataset despite them no longer being included in the analysis. The excluded studies are defined in this chunk of code. We have excluded this version the following studies:
-`r knitr::combine_words(exclude_from_analysis)`
Generally we label these as _old in the Google sheet but there are exceptions to this so we manually define them at the moment.

@@

my_bins <- c(0, 1, 5, 10, 20, 30, 40, 80)
mypalette <- colorBin(palette="YlOrBr", domain=my_spdf@data$Number, na.color="transparent", bins=my_bins)

@@

#Smoking prevalence by country
country_prevalence_list <- table_1 %>%
    group_by(country, .add = T) %>%
  mutate(., current_smok_percentage = current_smoker/total*100) %>%
  mutate(., former_smok_percentage = former_smoker/total*100) %>%
  mutate(., missing_percentage = missing/total*100) %>%
  mutate(., not_stated_percentage = not_stated/total*100) %>%
  mutate(never_smoker_percentage = never_smoker/total*100) %>%
  group_split()

@@

#save_as_docx(a, path = here('output_data', 'Table_1.docx'))
```

@@

included_studies <- quality_table_2 %>%
  filter(overall_quality != 'poor') %>%
  select(lead_author) %>%
  filter(!lead_author %in% c("reiter", "hippisley_cox", "ullah"))

meta <- tibble('author' = table_2$lead_author,
               'negative_smoker' = table_2$negative_current_smoker,
               'negative_never_smoker' = table_2$negative_never_smoker,
               'positive_smoker' = table_2$positive_current_smoker,
               'positive_never_smoker' = table_2$positive_never_smoker,
               'negative_former_smoker' = table_2$negative_former_smoker,
               'positive_former_smoker' = table_2$positive_former_smoker) %>%
        filter(author %in% included_studies$lead_author)
meta$author <- to_upper_camel_case(meta$author, sep_out = ", ")
meta$author <- plyr::mapvalues(meta$author,
                               from = cleaned_names,
                               to = correct_names)
study_dup <- filter(meta, author != "Niedzwiedz")
study_dup <- study_dup$author

testing_rr_current <- list()
for(study in 1:length(study_dup)){
  testing_rr_current[[study]] <- RR_testing(study_dup[study], "current")
}
testing_rr_former <- list()
for(study in 1:length(study_dup)){
  testing_rr_former[[study]] <- RR_testing(study_dup[study], "former")
}

data <- c(testing_rr_current, testing_rr_former)
k <- do.call(rbind.data.frame, data)

@@

#### Current versus never smoking for risk of severe disease
``` {r}
# Data --------------------------------------------------------------------
meta <- tibble('author' = table_4$lead_author,
               'non_severe_smoker' = table_4$non_severe_current_smoker,
               'non_severe_never_smoker' = table_4$non_severe_never_smoker,
               'severe_smoker' = table_4$severe_disease_current_smoker,
               'severe_never_smoker' = table_4$severe_disease_never_smoker,
               'non_severe_former_smoker' = table_4$non_severe_former_smoker,
               'severe_former_smoker' = table_4$severe_disease_former_smoker) %>%
  filter(author %in% included_studies$lead_author) %>%
  filter(author != "hippisley_cox")
